\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-12-05 archive.tex
spotlight before Armando revisions

\begin{table*}%[t!]
  \makebox[\textwidth][c]{
    \scriptsize
  \tabcolsep=4pt
  \renewcommand\code\texttt
  \renewcommand\codechar[1]{\texttt{"#1"}}
  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{>{\hspace{-0em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
    \toprule
    &{\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=7pt
      \begin{tabular}{cc}
        \begin{tabular}{c}
          \code{[7\, 2\, 3]}$\to $\code{[7\, 3]}         \\
          \code{[1\, 2\, 3\, 4]}$\to $\code{[3\, 4]} \\
          \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
          \\
          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
        \end{tabular}
        &
        \hspace{-0.3cm}\begin{tabular}{c}
          \code{[7\, 3]}$\to $\code{False}                              \\
          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
        \end{tabular}
      \end{tabular}
    }
    &
    \hspace{-0.5cm}\begin{tabular}{c}
      +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
      +83 973-831$\to $83.973.831\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{-}
      }}\\
      \hspace{1.25cm}\pop{\code{($f_0$ \codechar{.} \codechar{ }}}\\
      \hspace{1.4cm}\pop{\code{(cdr s)))}}\\
      ~\\
      Temple Anna H $\to $TAH\\
      Lara Gregori$\to $LG\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}\\
    \end{tabular}
    &
    \begin{tabular}{cc}
      \includegraphics[width = 3em]{figures/functions/4.png}&
      \includegraphics[width = 3em]{figures/functions/146}\\
      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
      ~\\
      \includegraphics[width = 3em]{figures/functions/112.png}&
        \includegraphics[width = 3em]{figures/functions/92.png}
      \\
      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\

@@

\noindent\textbf{Sleep-R: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$Q(p|x)$ to assign high probability to programs $p$ where
$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because incorporating those programs
into the frontiers will most increase $\lowerBound$.

Intuitively, this 3-phase inference procedure can work in practice because each of the 3 phases bootstraps off of the others (Figure~\ref{feeding}). As the DSL grows and as the recognition model becomes more accurate, waking becomes more effective, allowing the agent to solve more tasks; when we solve more tasks during waking,
the Sleep-G phase has more data from which to learn the DSL; and, because Sleep-R is trained on both
samples from the DSL and programs found during waking,
the recognition model gets both more data, and higher-quality data, whenever the DSL improves and whenever we discover more successful programs.
\begin{figure}\centering
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};

@@

Formally, $Q$ should approximate the true posteriors over programs.
We can either train $Q$ to perform full posterior inference by minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|Q(p|x) \right) \right]$,
or we can train $Q$ to perform MAP inference
by maximizing $\expect\left[\max_{p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]} \log Q(p|x) \right]$,
where in both cases the expectation is taken over tasks. Taking this expectation over the empirical distribution of tasks trains $Q$ on the real data; taking it over samples from the generative model trains $Q$ on ``dreams.''
We define a pair of alternative objectives for the recognition model,
$\mathcal{L}_{\text{RM}}^{\text{posterior}}$ and $\mathcal{L}_{\text{RM}}^{\text{MAP}}$,
which either train $Q$ to perform full posterior inference or MAP inference, respectively.
These objectives combine real-data and dream data:
\begin{align*}
  \mathcal{L}_{\text{RM}}^{\text{posterior}} &= \mathcal{L}_{\text{Real}}^{\text{posterior}} + \mathcal{L}_{\text{Dream}}^{\text{posterior}}&
  \mathcal{L}_{\text{RM}}^{\text{MAP}} &= \mathcal{L}_{\text{Real}}^{\text{MAP}} + \mathcal{L}_{\text{Dream}}^{\text{MAP}}\\
  \mathcal{L}_{\text{Real}}^{\text{posterior}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
    \frac{\probability\left[x,p|\mathcal{D},\theta \right]\log Q(p|x)}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\right] &
\mathcal{L}_{\text{Real}}^{\text{MAP}}& = \expect_{x\sim X}\left[\max_{\substack{p\in \mathcal{F}_x\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}} \log Q(p|x) \right]  \\
  \mathcal{L}_{\text{Dream}}^{\text{posterior}} &= \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log Q(p|x)\right]&
    \mathcal{L}_{\text{Dream}}^{\text{MAP}} &= \expect_{x\sim(\mathcal{D},\theta) }\left[\max_{\substack{p\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}}\log Q(p)\right]
\end{align*}
  The `dream' objectives are essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $q$ on those samples.
But, evaluating $\mathcal{L}_{\text{Dream}}$ involves drawing programs from
the current DSL, running them to get their outputs,
and then training $Q$ to regress from the input/outputs to the program.
Since these programs map inputs to outputs,
we need to sample the inputs as well.
Our solution is to sample the inputs
from the empirical observed distribution of inputs in $X$.

@@

    %\rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&
    \begin{tabular}{l}
\code{[2\, 1\, 4]$\to$[2\, 1\, 4\, 0]}\\
%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
\code{[9\, 8]$\to$[9\, 8\, 0]}\\
\blueCode{$f(\ell) = $($f_2$ cons $\ell$ (cons 0 nil))}\\\\

@@

\code{[4\, 2\, 6\, 4]$\to$[8\, 4\, 12\, 8]}\\
%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
\code{[2\, 3\, 0\, 7]$\to$[4\, 6\, 0\, 14]}\\
\blueCode{$f(\ell) = $($f_3$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\

      \code{[1\, 5\, 2\, 9]}$\to$\code{[1\, 2]}\\
      \code{[3\, 8\, 1\, 3\, 1\, 2]}$\to$\code{[3\, 1\, 1]}\\
      \blueCode{$f(\ell) = $($f_0$ empty? car}\\
      \phantom{\blueCode{$f(\ell) = $($f_0$}}\blueCode{($\lambda$ (l) (cdr (cdr l))) $\ell$)}
      \\

@@

%    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
    \begin{tabular}{l}
        \popp{\code{$f_0($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_0$ (n x))))}}\\
        \hspace{\helpSize}($f_0$: \emph{unfold})\\
        \popp{\code{$f_1($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if}}}\popp{\code{($f_1$ (- i 1) (cdr l))))}}\\
        \hspace{\helpSize}($f_1$: \emph{index})\\
        \popp{\code{$f_2($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
      \phantom{\code{$f_2($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_2$ (cdr l))))}}\\
        \hspace{\helpSize}($f_2$: \emph{fold})\\
        \greenCode{$f_3($f$,$l$)\,=\,$($f_2$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
        \hspace{\helpSize}($f_3$: \emph{map})\\
        \greenCode{$f_4(\ell)$\,=\,(if (empty? $\ell$) 0 (+ 1 ($f_4$ (cdr $\ell$)))))}\\
        \hspace{\helpSize}($f_4$: \emph{length})\\
        \greenCode{$f_5(n)$\,=\,($f_0$ (= $n$) ($\lambda$ (x) x) (+ 1) 0)}\\
        \hspace{\helpSize}($f_5$: \emph{range})

@@

\subsection{Refactoring code with version spaces}\label{appendixVersion}
Recall that our goal is to define an operator over version spaces, $I\beta_n$,
which calculates the set of $n$-step refactorings.
We define this operator recursively as follows:
\begin{align*}
  I\beta_n(v)& = \uplus \left\{ \underbrace{I\beta'(I\beta'(I\beta'(\cdots}_{i \text{ times}} v))) \;:\; 0\leq i \leq n \right\}\\
  I\beta'(v)& = \uplus \left\{(\lambda b)v\;:\;v\mapsto b\in S_0(v)\text{, when }v\not=\Lambda \right\}\cup
  \begin{cases}
    \text{if $v$ is a primitive or index:}&\varnothing\\
    \text{if $v = \lambda b$:}&\lambda I\beta'(b)\\
    \text{if $v = (f\;x)$:}&(I\beta'(f)\;I\beta'(x))\\
  \end{cases}
  \\
  S_k(v)& = \left\{\downshift{k}_0v\mapsto \$k \right\}\cup
  \begin{cases}
    \text{if $v$ is primitive:}&\left\{\Lambda\mapsto v \right\}\\    
    \text{if $v = \$i$ and $i < k$:}&\left\{\Lambda\mapsto \$i \right\}\\
    \text{if $v = \$i$ and $i\geq k$:}&\left\{\Lambda\mapsto \$(i + 1) \right\}\\
    \text{if $v = \lambda b$:}&\left\{v'\mapsto \lambda b' \;:\; v'\mapsto b'\in S_{k + 1}(b)  \right\}\\
    \text{if $v = (f\;x)$:}&\left\{v_1\cap v_2\mapsto (f'\;x') \;:\; v_1\mapsto f'\in S_k(f),\; v_2\mapsto x'\in S_k(x) \right\}\\
    \text{if $v = \uplus V$:}&\bigcup_{v'\in V}S_n(v')\\
    \text{if $v$ is $\varnothing$:}&\varnothing\\
    \text{if $v$ is $\Lambda$:}&\left\{\Lambda\mapsto\Lambda \right\}
  \end{cases}\\
  \downshift{k}_c \$i& = \$i\text{, when $i < c$}\\
  \downshift{k}_c \$i& = \$(i - k)\text{, when $i\geq c + k$}\\
  \downshift{k}_c \$i& = \varnothing \text{, when $c\leq i <  c + k$}\\
  \downshift{k}_c \lambda b &= \lambda\downshift{k}_{c + 1}b\\
  \downshift{k}_c (f\;x)& = (\downshift{k}_cf\;\downshift{k}_cx)\\
  \downshift{k}_c \uplus V& = \uplus \left\{\downshift{k}_c v \;|\;v\in V \right\}\\
  \downshift{k}_c v& = v\text{, when }v\text{ is a primitive or }\varnothing \text{ or }\Lambda
\end{align*}
where $\shift{k}$ is the shifting operator~\cite{pierce},
which adds $k$ to all of the free variables in a $\lambda$-expression or version space.
We have written the definition of $I\beta_n$ recursively,
but implement it using a dynamic program:
we hash cons each version space,
and only calculate the operators $I\beta_n$,
$I\beta'$, and $S_k$ once per each version space.
Refactoring is similarly done more quickly with dynamic programming (see Equation~\ref{FACTORING} for the recursive definition of \textsc{refactor}).

@@

>>>>

\begin{tabular}{}
  \end{tabular}

\begin{table*}%[t!]
  \makebox[\textwidth][c]{
    \scriptsize
  \tabcolsep=4pt
  \renewcommand\code\texttt
  \renewcommand\codechar[1]{\texttt{"#1"}}
  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{>{\hspace{-0em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
    \toprule
    &{\normalsize List Processing}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=7pt
      \begin{tabular}{cc}
        \begin{tabular}{c}
          \code{[7\, 2\, 3]}$\to $\code{[7\, 3]}         \\
          \code{[1\, 2\, 3\, 4]}$\to $\code{[3\, 4]} \\
          \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
          \\
          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
        \end{tabular}
        &
        \hspace{-0.3cm}\begin{tabular}{c}
          \code{[7\, 3]}$\to $\code{False}                              \\
          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
        \end{tabular}
      \end{tabular}
    }
    &
    \hspace{-0.5cm}\begin{tabular}{c}
      +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
      +83 973-831$\to $83.973.831\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{-}
      }}\\
      \hspace{1.25cm}\pop{\code{($f_0$ \codechar{.} \codechar{ }}}\\
      \hspace{1.4cm}\pop{\code{(cdr s)))}}\\
      ~\\
      Temple Anna H $\to $TAH\\
      Lara Gregori$\to $LG\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}\\
    \end{tabular}
    &
    \begin{tabular}{cc}
      \includegraphics[width = 3em]{figures/functions/4.png}&
      \includegraphics[width = 3em]{figures/functions/146}\\
      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
      ~\\
      \includegraphics[width = 3em]{figures/functions/112.png}&
        \includegraphics[width = 3em]{figures/functions/92.png}
      \\
      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\

@@

\noindent\textbf{Sleep-R: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$Q(p|x)$ to assign high probability to programs $p$ where
$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because incorporating those programs
into the frontiers will most increase $\lowerBound$.
This neural net is trained on two sources of data: samples, or `dreams',
drawn from the DSL,
as well as
replays of (program, task) pairs discovered during waking.

Intuitively, this 3-phase inference procedure can work in practice because each of the 3 phases bootstraps off of the others (Figures~\ref{feeding}~and~\ref{bayesianSummary}). As the DSL grows and as the recognition model becomes more accurate, waking becomes more effective, allowing the agent to solve more tasks; when we solve more tasks during waking,
the Sleep-G phase has more data from which to learn the DSL; and, because Sleep-R is trained on both
samples from the DSL and programs found during waking,
the recognition model gets both more data, and higher-quality data, whenever the DSL improves and whenever we discover more successful programs.
\begin{figure}\centering
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs: $p$};
    \node[align = center](DSL) at (3,1) {DSL: $\mathcal{D}$};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model: $Q$};

@@

Formally, $Q$ should approximate the true posteriors over programs.
We can either train $Q$ to perform full posterior inference by minimizing the expected KL-divergence, $  \expect\left[\text{KL}\left(\probability[p|x,\mathcal{D},\theta]\|Q(p|x) \right) \right]$,
or we can train $Q$ to perform MAP inference
by maximizing $\expect\left[\max_{p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]} \log Q(p|x) \right]$,
where in both cases the expectation is taken over tasks. Taking this expectation over the empirical distribution of tasks trains $Q$ on the real data; taking it over samples from the generative model trains $Q$ on ``dreams.''
We define a pair of alternative objectives for the recognition model,
$\mathcal{L}_{\text{RM}}^{\text{posterior}}$ and $\mathcal{L}_{\text{RM}}^{\text{MAP}}$,
which either train $Q$ to perform full posterior inference or MAP inference, respectively.
These objectives combine real-data and dream data:
\begin{align*}
  \mathcal{L}_{\text{RM}}^{\text{posterior}} &= \mathcal{L}_{\text{Real}}^{\text{posterior}} + \mathcal{L}_{\text{Dream}}^{\text{posterior}}&
  \mathcal{L}_{\text{RM}}^{\text{MAP}} &= \mathcal{L}_{\text{Real}}^{\text{MAP}} + \mathcal{L}_{\text{Dream}}^{\text{MAP}}\\
  \mathcal{L}_{\text{Real}}^{\text{posterior}}& = \expect_{x\sim X}\left[\sum_{p\in \mathcal{F}_x}
    \frac{\probability\left[x,p|\mathcal{D},\theta \right]\log Q(p|x)}{\sum_{p'\in \mathcal{F}_x}\probability\left[x,p'|\mathcal{D},\theta \right]}\right] &
\mathcal{L}_{\text{Real}}^{\text{MAP}}& = \expect_{x\sim X}\left[\max_{\substack{p\in \mathcal{F}_x\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}} \log Q(p|x) \right]  \\
  \mathcal{L}_{\text{Dream}}^{\text{posterior}} &= \expect_{(p,x)\sim(\mathcal{D},\theta) }\left[\log Q(p|x)\right]&
    \mathcal{L}_{\text{Dream}}^{\text{MAP}} &= \expect_{x\sim(\mathcal{D},\theta) }\left[\max_{\substack{p\\p\text{ maxing }\probability[\cdot |x,\mathcal{D},\theta]}}\log Q(p)\right]
\end{align*}
  The `dream' objectives are essential for data efficiency:
all of our experiments train \system on only a few hundred tasks, which is too little for
a high-capacity neural network.
Once we bootstrap a $(\mathcal{D},\theta)$,
we can draw unlimited samples from $(\mathcal{D},\theta)$
and train $Q$ on those samples.
But, evaluating $\mathcal{L}_{\text{Dream}}$ involves drawing programs from
the current DSL, running them to get their outputs,
and then training $Q$ to regress from the input/outputs to the program.
Since these programs map inputs to outputs,
we need to sample the inputs as well.
Our solution is to sample the inputs
from the empirical observed distribution of inputs in $X$.

@@

    \begin{tabular}{l}
      \code{[1\, 9]}$\to $\code{2}\\
      \code{[5\, 3\, 8]$\to$3}\\
    \blueCode{$f(\ell) = $($f_4$ $\ell$)}\\\\
      \code{[true\, false]}$\to $\code{2}\\
      \code{[false\, false\, false]$\to$3}\\
    \blueCode{$f(\ell) = $($f_4$ $\ell$)}\\\\

\code{[2\, 1\, 4]$\to$[2\, 1\, 4\, 0]}\\
%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
\code{[9\, 8]$\to$[9\, 8\, 0]}\\
\blueCode{$f(\ell) = $($f_2$ cons $\ell$ (cons 0 nil))}\\\\

\code{[2\, 1\, 4]$\to$[2\, 1]}\\
%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
\code{[9\, 8]$\to$[9]}\\
\blueCode{$f(\ell) = $($f_0$ ($\lambda$ (z) (empty? (cdr z))) car cdr $\ell$)}\\\\

@@

\code{[4\, 2\, 6\, 4]$\to$[8\, 4\, 12\, 8]}\\
%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
\code{[2\, 3\, 0\, 7]$\to$[4\, 6\, 0\, 14]}\\
\blueCode{$f(\ell) = $($f_3$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\
\code{[4\, 2\, 6\, 4]$\to$[-4\, -2\, -6\, -4]}\\
%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
\code{[2\, 3\, 0\, 7]$\to$[-2\, -3\, -0\, -7]}\\
\blueCode{$f(\ell) = $($f_3$ (- 0) $\ell$)}\\\\
\code{[4\, 2\, 6\, 4]$\to$[5\, 3\, 7\, 5]}\\
%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
\code{[2\, 3\, 0\, 7]$\to$[3\, 4\, 1\, 8]}\\
\blueCode{$f(\ell) = $($f_3$ (+ 1) $\ell$)}\\\\

      \code{[1\, 5\, 2\, 9]}$\to$\code{[1\, 2]}\\
      \code{[3\, 8\, 1\, 3\, 1\, 2]}$\to$\code{[3\, 1\, 1]}\\
      \blueCode{$f(\ell) = $($f_0$ empty? car}\\
      \phantom{\blueCode{$f(\ell) = $($f_0$}}\blueCode{($\lambda$ (l) (cdr (cdr l))) $\ell$)}
      \\\\

      \code{3$\to $[0\, 1\, 2]}\\
      \code{2$\to $[0\, 1]}\\
      \blueCode{$f(n) = $($f_5$ ($\lambda$ (x) x) 0 $n$)}\\\\

      \code{3$\to $[0\, 1\, 2\, 3]}\\
      \code{2$\to $[0\, 1\, 2]}\\
      \blueCode{$f(n) = $($f_5$ ($\lambda$ (x) x) 0 (+ 1 $n$))}\\\\

@@

%    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
    \begin{tabular}{l}
        \popp{\code{$f_0($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_0$ (n x))))}}\\
        \hspace{\helpSize}($f_0$: \emph{unfold})\\
        \popp{\code{$f_1($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if}}}\popp{\code{($f_1$ (- i 1) (cdr l))))}}\\
        \hspace{\helpSize}($f_1$: \emph{index})\\
        \popp{\code{$f_2($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
      \phantom{\code{$f_2($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_2$ (cdr l))))}}\\
        \hspace{\helpSize}($f_2$: \emph{fold})\\
        \greenCode{$f_3($f$,$l$)\,=\,$($f_2$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
        \hspace{\helpSize}($f_3$: \emph{map})\\
        \greenCode{$f_4(\ell)$\,=\,(if (empty? $\ell$) 0 (+ 1 ($f_4$ (cdr $\ell$)))))}\\
        \hspace{\helpSize}($f_4$: \emph{length})\\
        \greenCode{$f_5(\code{f},\code{m},\code{n})$\,=\,($f_0$ (= m) f (+ 1) n)}\\
        \hspace{\helpSize}($f_5$: \emph{generalization of range})

@@

\subsection{Refactoring code with version spaces}\label{appendixVersion}
Recall that our goal is to define an operator over version spaces, $I\beta_n$,
which calculates the set of $n$-step refactorings.
We define this operator recursively in terms of another operator, $I\beta'$, which performs a single step of refactoring:
\begin{align*}
  I\beta_n(v)& = \uplus \left\{ \underbrace{I\beta'(I\beta'(I\beta'(\cdots}_{i \text{ times}} v))) \;:\; 0\leq i \leq n \right\}
\end{align*}
where
  \begin{align*}
    I\beta'(u)& = \uplus \left\{(\lambda b)v\;:\;v\mapsto b\in S_0(u)%\text{, when }v\not=\Lambda
    \right\}\cup
  \begin{cases}
    \text{if $u$ is a primitive or index or $\varnothing $ or $\Lambda$:}&\varnothing\\
    \text{if $u = \lambda b$:}&\lambda I\beta'(b)\\
    \text{if $u = (f\;x)$:}&(I\beta'(f)\;I\beta'(x))\\
    \text{if $u = \uplus V$:}&\uplus\left\{I\beta'(u')\;|\;u'\in V \right\}
  \end{cases}  
  \end{align*}
  where we have defined $I\beta'$
  in terms of another operator, $S_k:\text{VS}\to 2^{\text{VS}\times\text{VS}}$,
  whose purpose is to construct the set of
  substitutions that are refactorings of a program in a version space.
We define $S$ as:
  \begin{align*}
  S_k(v)& = \left\{\downshift{k}_0v\mapsto \$k \right\}\cup
  \begin{cases}
    \text{if $v$ is primitive:}&\left\{\Lambda\mapsto v \right\}\\    
    \text{if $v = \$i$ and $i < k$:}&\left\{\Lambda\mapsto \$i \right\}\\
    \text{if $v = \$i$ and $i\geq k$:}&\left\{\Lambda\mapsto \$(i + 1) \right\}\\
    \text{if $v = \lambda b$:}&\left\{v'\mapsto \lambda b' \;:\; v'\mapsto b'\in S_{k + 1}(b)  \right\}\\
    \text{if $v = (f\;x)$:}&\left\{v_1\cap v_2\mapsto (f'\;x') \;:\; v_1\mapsto f'\in S_k(f),\; v_2\mapsto x'\in S_k(x) \right\}\\
    \text{if $v = \uplus V$:}&\bigcup_{v'\in V}S_n(v')\\
    \text{if $v$ is $\varnothing$:}&\varnothing\\
    \text{if $v$ is $\Lambda$:}&\left\{\Lambda\mapsto\Lambda \right\}
  \end{cases}\\
  \downshift{k}_c \$i& = \$i\text{, when $i < c$}\\
  \downshift{k}_c \$i& = \$(i - k)\text{, when $i\geq c + k$}\\
  \downshift{k}_c \$i& = \varnothing \text{, when $c\leq i <  c + k$}\\
  \downshift{k}_c \lambda b &= \lambda\downshift{k}_{c + 1}b\\
  \downshift{k}_c (f\;x)& = (\downshift{k}_cf\;\downshift{k}_cx)\\
  \downshift{k}_c \uplus V& = \uplus \left\{\downshift{k}_c v \;|\;v\in V \right\}\\
  \downshift{k}_c v& = v\text{, when }v\text{ is a primitive or }\varnothing \text{ or }\Lambda
\end{align*}
where $\shift{k}$ is the shifting operator~\cite{pierce},
which adds $k$ to all of the free variables in a $\lambda$-expression or version space,
and we have defined a new operator, $\downshift$, whose purpose is to
undo the action of $\shift$.
We have written definitions recursively,
but implement them using a dynamic program:
we hash cons each version space,
and only calculate the operators $I\beta_n$,
$I\beta'$, and $S_k$ once per each version space.
Refactoring is similarly done more quickly with dynamic programming (see Equation~\ref{FACTORING} for the recursive definition of \textsc{refactor}).

@@

From these results the consistency and completeness of $I\beta$ follows immediately by its definition.

%% First observe that
%% we do not \emph{enumerate} every refactoring,
%% but only those for which every expression of the form $(\lambda x.b)v$ has
%% $x$ free in $b$.
%% For example, we do not refactor $\code{+}$ into $(\lambda x. \code{+})v$ where $v$ is arbitrary,
%% because there are infinitely many such refactorings.
%% Define the following relation,
%% which captures this restricted kind of
%% refactoring:
%% \begin{gather*}
%%   \frac{f\reduce^\sim f'}{f\;x \reduce^\sim f'\;x}\qquad
%%   \frac{x\reduce^\sim x'}{f\;x \reduce^\sim f\;x'}\qquad
%%   \frac{b\reduce^\sim b'}{\lambda b \reduce^\sim \lambda b'}\qquad
%%   \frac{\$0 \text{ occurs free in }b}{(\lambda b)v\reduce^\sim \shift{-1}[\$0\mapsto \shift{1}v]b}
%%   \end{gather*}
%% \begin{theorem}
%%   \textbf{Completeness of $I\beta'$.}
%%   Let $v$ be a version space and let $p\in v$.
%%   If $p'\reduce^\sim p$
%%   then $p'\in I\beta'(v)$.
%% \end{theorem}
%% \begin{proof}
%%   We know (from the definition of $\reduce^\sim$) that $p'$ and $p$ differ only
%%   in that there exists a single subexpression in
%%   $p'$ of the form $(\lambda b)v$
%%   where that subexpression occurs in
%%   $p$ as $\shift{-1}[\$0\mapsto \shift{1}v]b$.
%%   From Theorem~\ref{}
%% \end{proof}
